\documentclass[12pt,english]{report}
\usepackage{mathptmx}
\renewcommand{\familydefault}{\rmdefault}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[a4paper]{geometry}
\setcounter{secnumdepth}{2} % Changed from 3 to 2. 0-chapter 1-section 2-subsection 
\setcounter{tocdepth}{2} % Changed from 3 to 2. 0-chapter 1-section 2-subsection 
\setlength{\parskip}{\medskipamount}
\setlength{\parindent}{0pt}
\usepackage{verbatim}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage{subfig} %% This package has to be here
\usepackage{setspace}
\usepackage{arabtex}
\usepackage[numbers]{natbib}
\usepackage{nomencl}
\usepackage{amsthm}
\usepackage{amsfonts}
\begin{document}

\chapter{Wavelets and the wavelet transform}

The topic of wavelet analysis is relatively a new concept. To fully comprehend the subject one usually needs to invest a large amount of time in studying the subject. In addition, it requires a high level of mathematical background, which is usually not covered in a basic calculus courses.
The Wavelet transform, like the Fourier transform, is used in many fields of mathematics, physics and engineering. Quite a few articles and books were written on the subject, therefore, here we attempt to provide a short and intuitive introduction to the topic and let the interested reader to learn more about the subject in the literature. The reader is referred to \cite{meyer1992wavelets} for a comprehensive discussion on the wavelet transform.\\ 

Recall that the Fourier Transform gives the frequency information of a signal, namely, it provide us information about the frequencies that can be found in the signal. However, it does not indicate on what time exactly these frequency components exist. In cases where the signal is stationary, this information is not needed. FT can also be used for non-stationary signals, if we are only interested in what frequency components exist in the signal, but are not required to know where in time these frequencies occur. Nevertheless, if we want to know what frequency component occurs at what time (interval), then Fourier transform is not the best transform to use. Wavelet transform provides us with the time-frequency representation of the signal and thus gives us information about the both the time and frequencies.\\

Wavelets may be seen as small waves $\psi(t)$, which oscillate few times, but unlike the harmonic waves ($sin(nx)$ and $cos(nx)$ on $(-\pi,\pi]$) must die out to zero as $t \rightarrow \pm\infty$. The most applicable wavelets are those that die out after a few oscillations on a finite interval $[a,b)$. \\

The wavelet transform aims to convert a signal into a series of wavelets and by doing so to provide a way for analyzing waveforms bounded in both frequency and duration. This is done to provide information that is not obvious in the original time domain function and also to allow signals to be stored efficiently and to be able to better approximate real-world signals. \\

\begin{equation}
\Psi^{\psi}_{x}(\tau, s)=\frac{1}{\sqrt{|s|}}\int{x(t)\psi^*(\frac{t-\tau}{s})dt}
\label{eq:wavelet_decomposition}
\end{equation}
As seen in the above Equation \ref{eq:wavelet_decomposition}, the transformed signal is a two variables function, $\tau$ and $s$ , the translation and scale parameters respectively. $\psi(t)$ is called the mother wavelet.\\

The scale parameter $s$ can be thought of as exactly the scaling in maps. High scales correspond to a non-detailed high level view (of the signal), and low scales correspond to a detailed low level view. Similarly, in terms of frequency, low frequencies (high scales) correspond to a global information of a signal (that usually spans the entire signal), whereas high frequencies (low scales) correspond to a detailed information of a hidden pattern in the signal (that usually lasts a relatively short time).\\

The idea of the wavelet transform is analyzing a signal function $f(x)$ by creating a mathematical structure that vary in scale. i.e., construct a function, shift it by some amount, change its scale. Repeat the procedure to obtain a new approximation. This is usually refereed to as a \emph{multiresolution analysis}. In order to analyze a non-stationary signal, we need to determine its behavior at any individual event. Multi-resolution analysis provides one means to do this. A multi-resolution analysis decomposes a signal into a smoothed version of the original signal and a set of detail information at different scales. Once we have decomposed a signal this way, we may analyze the behavior of the detail information across the different scales.\\

The fundamental idea of wavelet transforms is that the transformation should allow only changes in time extension, but not in the shape of the wavelet. 
A function $f$ in $\mathbb{R}$ can be expressed in terms of wavelet series as follows:
\begin{equation}
f(x)=\sum\limits_{k}{f_k \phi(x-k)}+\sum\limits_{k}{f_{j,k}\psi_{j,k}(x)}
\end{equation}

where $\phi$ is the scaling function and $\psi$ is the wavelet. $k$ runs through all integer and represents shifts.\\


\end{document}
